\section{Suggestions for improvement}
    One thing we have learned from this project, is that software development
    requires many iterations, and thus one have to beware not to approach
    the result from the design, analysis or implementation as final, but rather
    as a state of some running evaluation. The course material seems to agree
    with the quote "a project plan is developed iteratively"\cite{oose} p.698
    and also in \cite{mad-article} s.23 fig 7.

    Also we would have been able to improve the project process, if we had been
    better to set demands for deadline and information from Subsea7. Maybe we
    felt, that we ``needed'' Subsea7, more than they needed us, and thus we
    have been reluctant with demands. With more (and still fair) demands we
    would have been able to get a more comprehensive specification requirement,
    earlier in the process, which would have saved us some code. Also we would
    have been completely done at this point, if Subsea7 have had their xml-feed
    service ready in time. To hold our own deadlines we have been using a
    scrum-like organization model \cite{oose} p. 647, and we have had great
    success with this model. Thus we will suggest other devlopment teams to
    check out this model if they are not familiar with it.

    As mentioned earlier, our project is 98 pct. done and Subsea7 is pleased
    with the result, and so are we. Furthermore there have been only minor
    problems in the planing and development process. I think we did a good job,
    and looking back I don't see other things that we could have done (but not
    did), to improve the planning and the development process.
