\section{Parties involved in the project}
There have been two parties involved in the project, besides our development team. Below are a brief description of the parties:

\begin{description}
    \item [Subsea7 IT-department] 
        The IT-department has been our primary contact. All communication with 
        Subsea7 has been handled through this department. The department has 
        IT-knowledge which allowed discussions about the application to happen 
        in technical terms.   

    \item [End users at Subsea7]
        The contact with the end users, has been handled by the IT-department. 
        The only direct contact we have had with the end users has been as 
        part of think-out-laud experiments and similar tests.
\end{description}

The fact that Subsea7 has an quite big  IT-deprtment has had a big impact on the project. 
The it department had already made some decisions regarding the application 
before we entered the project, and at times it had been
difficult to bring new ideas to the table. 

As an example they did not see any requirement for the think-out-laud experiments.
They had already discussed how they wanted the application, and in their eyes, our
job was more or less 'just' to code it, where we wanted to discuss and challenge the decisions they had made in advance. Thus, there have been some discussions, where we needed
to explain that it was a requirement, that we could do the experiments
and present our suggestions for improvements. Otherwise we would not be able to 
produce good subreports for the course.

Besides the matter mentioned above, all collaboration has been more or less flawless.
