\section{Project collaboration}

\subsection{Meetings with client}

We have had 4 meetings with our client. The first was a video-meeting
the \nth{23} of February, here we was introduced to the client and the
project. The rest of the meetings took place at their office in
Stavanger, Norway.

The second meeting was \nth{21} of March, where we learned more about
the project in details. Discussed the UI paper prototypes we had made,
and some broad implementation details -- like that they would provide
a service serving XML. We borrowed two iPads from them to test the
application on.

Third meeting the \nth{18} of April, we had our first running prototype
running on the iPad covering most of the basic functionality. We
discussed the UI so far with the client, which gave reason to several
changes. We also did a user test, which showed that several things had
to be changed. We got a bit more into the technical details of the
iPads requesting data from their servers, talking about the XSD for
the XML data, and what fields should be in the configuration file.

The last meeting took place on the \nth{23} and \nth{24} of May, where
we worked at their offices. They where not completely finished with
their service providing data, so we had to wait to the second day
before we could really test S7Finder. We could perform some user tests
of the UI, and make some changes to it. We also realised some of the
functionality for updating data would never be used, so we could
remove a lot of complexity from our code, and simplified the XML
format used for data transfer.

\subsection{Communication with client}

Besides the meetings with the client, we've had some email exchanges
with them. Subsea7 has provided fast replies on all of them (even on
holidays), and generally been very responsive.

\subsection{Organising work}

We've had a very SCRUM-like approach to planning our work,
prioritising the remaining tasks and deciding which things should be
done the coming week -- always prioritising basic functionality
first. Most of the time during the PKSU course we have had two weekly
meetings, but the content of these changed over time as described in the following paragraphs. We've also had our own IRC channel where we've been able to
talk with each other when not at the university.

Up until the second meeting with the client we used the two weekly
meetings to discuss the current state of the project and the upcoming
report. We spent a lot of time talking about what to do, and how we
should do it, but not that much on actual work. For each meeting, a
set of meeting minutes were written.

After the second meeting in Stavanger, the development process became
much more fruitful, with a realisation that a we had to provide a
working preview at the third meeting in April. One of the weekly
meetings was dropped in favour of doing real work, and in the period
up to the third meeting we had an intense development process,
including daily Skype meetings to keep track of the current
progress. No meeting minutes have been written in this period, as the
meetings took on a more conversational manner.

After the third meeting with Subsea7, our two weekly meetings mostly
dealt with development. We started to use Pivotal Tracker to keep
track of the pending tasks, how far each team member was with his part
of the project, and prioritising future tasks. Instead of writing
meeting minutes, decisions would be entered into the Pivotal Tracker
-- this worked far better than writing resumes that would never be
read.
