\section{Team collaboration}
The internal team work in the group has been working remarkably well. We have been
working quite homogeneously and are able - as a team - to increase the work pace towards
deadlines and presentations. As an example we introduced daily Skype meetings
the week before our last presentation in Norway, where we discussed the
development progress, technical implementations and each team member presented his
personal work progress. This enabled us to keep a high pace, and enabled the
group to immediately identify problems in the work flow. We have also established
an IRC channel, where all members are present (mostly 24/7). Thus, it is easy
get feedback on implementation ideas etc., doing the day.

At that start of the project we decided to use Pivotal for project management,
but because the tasks and architecture were not that complex yet, it seemed
as if we did not benefit enough from the tool if we took into account the time
required to keep Pivotal up to date. And thus, we forgot/stopped using it over time.
As the project has grown we have found that it is hard to sustain overview of
the tasks that need to be completed. To restore a clear overview we have
agreed that every team member will use Pivotal for the rest of the development
process.

The only strenuous situation we have faced was in regard to the hand in of the
second sub-report. At that time, due to other exams, the communication was not
good enough. Thus there was not enough clarity in regard to where we were
in the progress of writing the sub-report. To face such situations in the future
we did some ongoing evaluation of each others work towards the
deadline, and at the weekly meetings we used the report requirements to
check that we had remembed all the required parts.



